Who does big team science?

### 2023-08-16 appendix.Rmd
making sure the figures are right
fixed aspect ratio

```{r fig-keywords, include = TRUE, fig.cap="Keyword Analysis for Each of the Four Subject Areas."}
health_key <- fig("figure/health_key.png")
physical_key <- fig("figure/physical_key.png") 
life_key <- fig("figure/life_key.png") 
social_key <- fig("figure/social_key.png")

@@

ggsave("figure/figure_app_1.png")
```

@@

```{r fig-inst, include = TRUE, fig.cap="Number of unique institutions involved in big-team science papers across decades."}
DF$inst_time$normalized <- DF$inst_time$unique / DF$inst_time$total * 100

@@

Types of publications are presented in Figure \@ref(fig:fig-pub-types). The patterns of publications are roughly similar for big team science authors and all authors. It appears that porportionally, big team members are more likely to post preprints in comparison to all authors.

```{r fig-pub-types, include = TRUE, fig.cap = "Types of publications for big-team science and all authors."}
pub_levels <- c("ab", "ar", "ip", "bk", "bz", "ch", "cp", 
                "cr", "dp", "ed", "er", "le", "no", "pp", 
                "pr", "rp", "re", "sh", "tb")
pub_labels <- c("Abstract", "Article", "In Press", "Book", "Business", "Book Chapter", "Conference Paper", 
                "Conference Report", "Data Paper", "Editorial", "Erratum", "Letter", "Note", 
                "Preprint", "Press Release", "Report", "Review", "Survey", "Retracted")

@@

ggplot(pub_types, aes(area = count, fill = binned_count,
               label = publication_type)) +
  geom_treemap() +
  facet_wrap(~type) + 
  geom_treemap_text(colour = "black", place = "centre",
                    size = 15, grow = FALSE) +  
  # scale_fill_brewer(name = "Sample Size", palette = "Spectral") +
  theme_classic()

ggsave("figure/figure_app_3.png")
# can probably figure out how to combine these feasibly to make it readable
# ab Abstract Report, ar Article, ip Article in Press, bk Book, bz Business Article, ch Book Chapter, cp Conference Paper, cr Conference Review, dp Data paper, ed Editorial, er Erratum, le Letter, no Note, pp Preprint, pr Press Release, rp Report, re Review, sh Survey, tb Retracted.
```

>>>>

```{r fig-keywords, include = TRUE, fig.cap="Keyword Analysis for Each of the Four Subject Areas.", fig.width=10, fig.height=6}
health_key <- fig("figure/health_key.png")
physical_key <- fig("figure/physical_key.png") 
life_key <- fig("figure/life_key.png") 
social_key <- fig("figure/social_key.png")

@@

ggsave("figure/figure_app_1.png", width = 10, height = 6)
```

@@

```{r fig-inst, include = TRUE, fig.cap="Number of unique institutions involved in big-team science papers across decades.", fig.width=10, fig.height=6}
DF$inst_time$normalized <- DF$inst_time$unique / DF$inst_time$total * 100

@@

Types of publications are presented in Figure \@ref(fig:fig-pub-types). The patterns of publications are roughly similar for big team science authors and all authors. It appears that proportionally, big team members are more likely to post preprints in comparison to all authors.

```{r fig-pub-types, include = TRUE, fig.cap = "Types of publications for big-team science and all authors.", fig.width=10, fig.height=6}
pub_levels <- c("ab", "ar", "ip", "bk", "bz", "ch", "cp", 
                "cr", "dp", "ed", "er", "le", "no", "pp", 
                "pr", "rp", "re", "sh", "tb")
pub_labels <- c("Abstract", "Article", "In Press", "Book", "Business", "Book Chapter", "Conference Paper", 
                "Conference Report", "Data Paper", "Editorial", "Erratum", "Letter", "Note", 
                "Preprint", "Press Release", "Report", "Review", "Survey", "Retracted")

@@

ggplot(pub_types, aes(area = count, fill = binned_count,
               label = publication_type)) +
  geom_treemap() +
  facet_wrap(~type) + 
  geom_treemap_text(colour = "black", place = "centre",
                    size = 15, grow = FALSE) +  
  scale_fill_brewer(name = "Sample Size", palette = "Spectral") +
  theme_classic()

ggsave("figure/figure_app_3.png", width = 10, height = 6)
# can probably figure out how to combine these feasibly to make it readable
# ab Abstract Report, ar Article, ip Article in Press, bk Book, bz Business Article, ch Book Chapter, cp Conference Paper, cr Conference Review, dp Data paper, ed Editorial, er Erratum, le Letter, no Note, pp Preprint, pr Press Release, rp Report, re Review, sh Survey, tb Retracted.
```

### 2023-08-16 manuscript_scopus.tex
fixed aspect ratio

\textbf{\emph{Education}}. We collected degree information from the author table.

\textbf{\emph{Types of Publications}}. We took information from the
publication type variable for each author's publications to present
information about the types of papers BTS authors publish.

\textbf{\emph{Publication Metrics}}. For each author, we calculated the total
number of publications, and the h-index. The h-index represents the
highest \emph{h} number of publications that have at least \emph{h} citations.

@@

Publication interest area was summarized by the four large subject areas creating a word cloud plot of the total number of publications within the ASJCs. Figure @\ref(fig:fig-clouds) displays that the health sciences tends to publish within medicine and oncology, with a corresponding focus of cancer research and genetics for the life sciences. The physical sciences is mostly dominated by physics research, chemistry, and ecology. The BTS publications in the social sciences are mostly within psychology, education, and health.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-career-1.pdf}
\caption{\label{fig:fig-career}Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year.}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  We will report descriptive statistics on the total number of
  publications and h-index for individuals overall.

  \begin{itemize}
  \tightlist
  \item
    Do this for each unique person and report averages
  \item
    Do this for each publication and create an average for each
    publication
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Next, we will use the same analyses described in the career length
  section to analyze trends over time. An increasing slope over time
  indicates that individuals who are publishing more are more
  represented in BTS over time (i.e., increasing numbers of scholars
  with higher publication rates), while a negative slope indicates
  more researchers with less publications.
\item
  A positive slope for standard deviation indicates increasing
  variance over time (i.e., more diversity in the individual
  publication rates), while a negative slope would indicate less
  diversity in researchers over time. While publication rates do not
  represent value as a researcher, they are often used in hiring and
  promotion decisions, and we will use this variable as a proxy to
  gauge the diversity in scholars represented in big teams.
\item
  We will separate these by each of the four subject areas.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  We will present visualizations of the country information listed for
  authors, and we will discuss the areas of world in which authors
  generally come from, as well as the lowest representation of
  authors.
\end{itemize}

@@

\begin{itemize}
\item
  To understand the change in representation diversity, we will
  summarize the total number of geopolitical regions for each paper.
  Using a linear model, we will examine if the number of regions
  present is predicted by the year of publication. Increasing
  diversity would be represented by a positive slope, while decreasing
  diversity would be represented by a negative slope.
\item
  Last, we will examine the differences in representation for
  corresponding author sets versus all other authors. For papers with
  10 to 49 authors, we will use the three first authors and the last
  author to compare against other authors. For 50 to 99 authors, five
  first authors plus last will be used, and for all papers with more
  than 100 authors, we will use ten first authors and the last author.
  We will calculate the frequencies of each of the UN Sub-Regions for
  first authors versus other authors, converting these values to
  proportions. Given the expected small sample sizes of these
  contingency tables, we will group together titles based on the year
  of publication (assuming at least 5 publications per year, these may
  be binned by 5-year or smaller increments to increase sample size).
  For each grouping, we will calculate the effect size of the
  differences in frequencies comparing first authors to all other
  authors. Since this data is categorical, we will use Cramer's \emph{V} to
  represent the effect size. If the effect size includes zero in its
  confidence interval, this result will imply that first and all other
  authors represent the same pattern of UN Sub-Region diversity. Any
  confidence interval that does include zero represents a difference
  in diversity.
\end{itemize}

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/author-gpe-figure-1.pdf}
\caption{\label{fig:author-gpe-figure}A comparison of author affiliation geopolitical region across decades. F stands for first authors and O stands for other authors.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/effect-gpe-figure-1.pdf}
\caption{\label{fig:effect-gpe-figure}Effect size of the differences in representation for UN Regions for author affiliations in big-team science papers by year.}
\end{figure}

@@

@@

@@

>>>>

\textbf{\emph{Education}}. We collected degree information from the author table. Information on this variable is in the appendix.

\textbf{\emph{Types of Publications}}. We took information from the
publication type variable for each author's publications to present
information about the types of papers BTS authors publish. Information on this variable is in the appendix.

\textbf{\emph{Publication Metrics}}. For each author, we calculated the total
number of publications, and the h-index. The h-index represents the
highest \emph{h} number of publications that have at least \emph{h} citations. \emph{h}-count was only used for descriptive statistics.

@@

Publication interest area was summarized by the four large subject areas creating a word cloud plot of the total number of publications within the ASJCs. Figure \ref{fig:fig-clouds} displays that the health sciences tends to publish within medicine and oncology, with a corresponding focus of cancer research and genetics for the life sciences. The physical sciences is mostly dominated by physics research, chemistry, and ecology. The BTS publications in the social sciences are mostly within psychology, education, and health.

@@

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-career-1.pdf}
\caption{\label{fig:fig-career}Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph.}
\end{figure}

@@

@@

We used the same analyses described in the career length section to analyze trends over time. An increasing slope over time indicates that individuals who are publishing more are more represented in BTS over time (i.e., increasing numbers of scholars with higher publication rates), while a negative slope indicates more researchers with less publications. A positive slope for the standard deviation of publication metrics indicates increasing variance over time (i.e., more diversity in the individual publication rates), while a negative slope would indicate less diversity in researchers over time. While publication rates do not represent value as a researcher, they are often used in hiring and promotion decisions, and we used this variable as a proxy to gauge the diversity in scholars represented in big teams. As shown in Figure \ref{fig:fig-heatmap} publication metrics were generally negative for the average publication metrics, indicating more scholars over time with lower numbers of citations with the strongest effects in health and social sciences. The variability of publication counts was not significant for the life sciences but was negative for the physical sciences (less variability over time) and positive for social and health sciences (more variability and over time). This result indicates that the physical sciences are trending toward scholars with less publications but also less diverse in number of publications, while the health and social sciences see more diversity in publication counts and less published scholars overall.

@@

@@

Author geopoligical region is displayed in Figure \ref{fig:fig-map-both}. Big team publications appear to be lead by North America and Western Europe, while all publications are lead by North America and East Asia. To understand the change in representation diversity, we examined if the number of regions in a publication is predicted by the year of publication. Increasing diversity would be represented by a positive slope, while decreasing diversity would be represented by a negative slope. As shown in Figure \ref{fig:fig-heatmap}, the physical sciences do not show a trend of change in representation, while all other sciences showed a positive effect increasing in the number of geopolitical regions authors represent on publications.

Last, we examined the differences in representation for corresponding author sets versus all other authors. For papers with 10 to 49 authors, we used the three first authors and the last author to compare against other authors. For 50 to 99 authors, five first authors plus last were used, and for all papers with more than 100 authors, we used ten first authors and the last author as the corresponding author set. We then calculated the frequencies of each of the UN Sub-Regions for corresponding authors versus all other authors, converting these values to proportions. Given the expected small sample sizes of these contingency tables, we grouped together titles based on the year of publication. For each grouping, we then calculated the effect size of the differences in frequencies comparing corresponding authors to all other authors. Since this data is categorical, we used Cramer's \emph{V} to represent the effect size. If the effect size includes zero in its confidence interval (to four decimal places), this result will imply that first and all other authors represent the same pattern of UN Sub-Region diversity. Any confidence interval that does include zero represents a difference in diversity.

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-author-gpe-1.pdf}
\caption{\label{fig:fig-author-gpe}A comparison of author affiliation geopolitical region across decades. F stands for first authors and O stands for other authors.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_scopus_files/figure-latex/fig-effect-gpe-1.pdf}
\caption{\label{fig:fig-effect-gpe}Effect size of the differences in representation for UN Regions for author affiliations in big-team science papers by year. Larger dots indicate more papers and authors represented in the calculation of effect size.}
\end{figure}

Figure \ref{fig:fig-author-gpe} indicates the percent of authors in regions. In general, we found the same pattern as the overall analysis wherein most authors are from Europe and North America. The pattern of representation is roughly similar for the separation of small, medium, and large numbers of authors on papers. Across time, the representation does appear to diversify, with more representation in Asia, Latin American, and Africa. Figure \ref{fig:fig-effect-gpe} represents the size of the differences in first/corresponding authors and other authors across time and number of authors. The differences in representation are larger for papers with more authors; however, the effects are non-zero for many of the comparisons. Encouragingly, over time these effects appear to diminish in size. One limitation with the calculation of effect sizes for count data is the sensitivity of the data to sample size (i.e., \(\chi^2\) is upwardly biased by sample size, and \(V\) is calculated based on this value). While we used the inclusion of zero as our boundary for ``significance'', the interpretation of the effects is that most are likely small: \(V\) \textless{} .05: 31.79\%, \textless{} .10: 70.20\%, \textless{} .20: 94.04\%.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\begin{itemize}
\tightlist
\item
  number of publications increasing
\item
  research areas appear to be cancer, physics, and psychology
\item
  the number of authors is increasing as well
\item
  career length is decreasing, number of publications necessary to be involved is decreasing
\item
  number of gpe increasing
\item
  appears to be slowly diversifying, yet still not equal in first and other authors in diversity
\end{itemize}

limitations - based on the data we could get, we made up the definition of BTS, under-representation of articles in other languages/that they don't collect

future: incentives for bts and why people do it

@@

\hypertarget{supplemental-materials}{%
\section{Supplemental Materials}\label{supplemental-materials}}

@@

\textbf{\emph{Education}}. As noted in our pre-registration, we would only present this variable if we could obtain at least 50\% information on the authors who publish in big team science papers. 95.83\% of the data was not available.

@@

Types of publications are presented in Figure \ref{fig:fig-pub-types}. The patterns of publications are roughly similar for big team science authors and all authors. It appears that porportionally, big team members are more likely to post preprints in comparison to all authors.
